### 2013-08-30 authors.tex
Гравный файл и файлы авторства

\bigskip

\noindent{\tolerance 300 \footnotesize
\textbf{Глазунов Вадим Валерьевич}~—  аспирант кафедры <<Телематика (при ЦНИИ РТК)>> ИПММ СПбГПУ. neweagle@gmail.com; СПбГПУ, Политехническая ул., д.~29, г.~Санкт-Петербург, 195251, Россия; р.т. +7(812)552-65-21. Научный руководитель~--- Курочкин~М.\,А.
\vskip 1mm
\noindent\textbf{Glazunov V. Vadim}~—  of . neweagle@gmail.com; SPbSPU , kiy pr., 29., St. Petersburg, 195251, Russia; office phone +7(812)552-6521, Scientific advisor~--- Kurochkin~M/\,A.
\vskip 1mm
\noindent\textbf{Курочкин Леноид Михайлович} — к.т.н., доцент кафедры <<Телематика (при ЦНИИ РТК)>> института прикладной математики и механики Санкт-Петербургского государственного Политехнического университета. Область научных интересов: . Число научных публикаций — . ALT@iias.spb.su, www.tulupyev.spb.ru; СПИИРАН, 14-я линия В.О., д. 39, г. Санкт-Петербург, 195251, РФ; р.т. +7(812)552-65-21.
\vskip 1mm
\noindent\textbf{Alexander L. Tulupyev} — PhD in Computer Science, Associate Professor; acting Head of Theoretical and Interdisciplinary Computer Science Laboratory, SPIIRAS, Associate Professor of Computer Science Department, SPbSU. Research area: uncertain data and knowledge representation and processing, mathematics and computer science applications in socio-cultural studies, biostatistics, simulation, and mathematical modeling applications in epidemiology, data intensive software systems development technology. Number of publications — 160. ALT@iias.spb.su, www.tulupyev.spb.ru; SPIIRAS, 14-th line V.O., 39, St. Petersburg, 199178, Russia; office phone +7(812)328-3337, fax +7(812)328-4450.\vskip 1mm}

{\footnotesize \noindent\textbf{Поддержка исследований.} Работа выполнена при финансовой поддержке РФФИ, проект № 13-07-12106 офи_м.
}